A Visibility-Based Combinatorial Data Structure for Mobile Robots in Polygons

### 2018-05-16 final_paper/paper.tex
mostly grammar edits

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage[font=footnotesize,labelfont=footnotesize]{caption}
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{subcaption}
\usepackage{hyperref} % use for hypertext links, including those to external documents and URLs
\usepackage{multirow}
\usepackage{rotating}
\usepackage{array}
\usepackage{xspace}
\usepackage{indentfirst}
\usepackage{amssymb}
\usepackage{float}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{comment}
\usepackage{sidecap}
\usepackage{booktabs}

\title{Mobile Robots in Polygons: a Combinatorial Data Structure for
Navigation, Localization, and Coverage}

@@

The \textit{vertex visibility graph} $G_V(P) = (V, E)$ has $V = \{v_0, v_1, \ldots, v_{n-1}\}$ and $v_iv_j\in E$ if and only if the segment connecting $v_i$ and $v_j$ is completely contained in the interior of $P$. We call $v_i, v_j$ visible to each other if $v_iv_j\in E$. By definition, the vertex visibility graph for a convex polygon with $n$ vertices is a $n-$clique. Fig~\ref{fig:vvg} shows the visibility graph for a non-convex polygon.
\begin{figure}
    \includegraphics[width=0.8\linewidth]{images/viz_graph.png}
    \centering
    \caption{The vertex visibility graph}\label{fig:vvg}
    \centering
\end{figure}

We can also define the \textit{vertex-edge visibility graph} for a polygon. Using the definition in \cite{rourke_viz}, the \textit{vertex-edge visibility graph} for $P$, $G_E(P) = (V, E)$, is a bipartite graph where $V = \{\text{all vertices and edges in $P$}\}$ and $v_ie_j\in E$ if and only if $v_i$ can see the interior of $e_j$, i.e., $\exists$ point $q$ on the open edge $e_j$ such that $v_i$ and $q$ are visible. \cite{rourke_viz} established that the vertex-edge visibility graph is equivalent to the edge-edge visibility graph. Fig~\ref{fig:veg} shows an example for vertex-edge visibility graph.
\begin{figure}
\centering
\subfigure[]{
\centering
  \includegraphics[width=0.6\linewidth]{images/concave_pent.png}
  \centering
  \label{fig:c_p}
}
\subfigure[]{
  \includegraphics[width=0.8\linewidth]{images/viz_edge_graph.png}
  \centering
}
\caption{The vertex-edge visibility graph for the polygon in Fig~\ref{fig:c_p}}\label{fig:veg}

@@

The partial local sequence for a vertex $v_0$ in polygon $P$ contains new vertices created by the intersection of a ray, $r$, from each of the other vertices in $P$ to $v_0$ and the boundary of $P$ if $r$ exits $P$ at an edge of $P$. For example, in Fig~\ref{fig:pls}, the vertices in the partial local sequence of $v_0$ but not in the original polygon is annotated with blue dots.
\begin{figure}
    \includegraphics[width=0.8\linewidth]{images/partial_local_sequence.png}
    \centering
    \caption{Partial local sequence for $v_0$}\label{fig:pls}
    \centering
\end{figure}
The partial local sequences for vertices in a polygon is helpful for defining the equivalent class of edge visibility.

@@

\section{Bounce Visibility Diagram}
We will now define the central structure for this paper, the \textit{Bounce Visibility Diagram} for polygon $P$. We will first insert all new vertices appeared in the partial local sequence for all vertices into the polygon $P$ to get the new polygon $P'$ with $m$ vertices. We intentionally create degeneracies here for reasons that will be revealed later.

The $x-$axis of the diagram parameterizes the boundary of the new polygon, denoted as $\partial P'$, in counterclockwise order to  $[0, m+1)$. We map every point in $\partial P'$ to a point in $[0, m+1)$; vertex $v_i$ on $\partial P'$ will map to $x = i$; all points in the open edge $v_iv_{i+1}$ will uniformly map to $(i, i+1)$. Let $f: [0, m+1)\mapsto \partial P'$ denote this parameterization.

The vertical axis $\theta$ of the diagram has range $[0, \pi]$, representing all the possible angles the robot can bounce to. At angle $0$, the robot can bounce to the vertex in the front when it travels counterclockwise on $\partial P'$. Let the robot's ``forward direction'' be the direction of the edge that the robot is on in counterclockwise order.

For each vertex $v_i$, we can define a function $B_i: [0, n) \mapsto [0, \pi]$ for the angle between the robot's forward direction and its direction to vertex  $v_i$ when the robot stands at $f(x)\in \partial P'$. $B_i$ are piecewise continuous functions; the discontinuities occur when $x$ is an integer, or, in other words, at the vertices of $\partial P'$. The original vertices in $P$ mark discontinuities because the angle reference frame for the robot changes when it moves to a new edge; the new vertices in $P'$ mark discontinuities because the robot has different vertex visibility view around those vertices, which motivates us to compute the partial local sequences in the first place. Fig~\ref{fig:bvd} shows an example of the bounce visibility diagram. 
\begin{figure}
\centering
\subfigure[]{
\centering
\includegraphics[width=0.7\linewidth]{images/color_pent.png}
\label{fig:color_pent}
}
\subfigure[]{
\centering
\includegraphics[width=0.8\linewidth]{images/bvd.png}
}
\caption{The bounce visibility diagram for the polygon shown in Fig~\ref{fig:color_pent}. }
\label{fig:bvd}

@@

\begin{figure}
% \makebox[\linewidth][c]{%
\subfigure[]{
\centering
\includegraphics[width=0.5\linewidth]{images/regular_pent.png}
}
\centering
\subfigure[]{
\centering
\includegraphics[width=0.8\linewidth]{images/regular_pent_bvd.png}
}
\centering
% }\\
\caption{The bounce visibility diagram for the regular pentagon. }
\label{fig:regular_pent_bvd}
\end{figure}
We can easily extract information for the fixed angle bouncing strategy mentioned in section~\ref{subsec:bounce_strategy} by overlaying an horizontal line $\theta = \theta_0$ onto the BVD, where $\theta_0$ is the angle chosen by the bouncing strategy. In Fig~\ref{fig:limit_cycle_bvd}, one of the possible limit cycle with fixed angle $\theta = 0.6$ for the concave quadrilateral is shown in Fig~\ref{fig:limit_cycle}; the bounce visibility diagram for this polygon is shown in Fig~\ref{fig:concave_quad_bvd} with the horizontal line $\theta = \theta_0$, where $\theta_0 = \frac{\pi}{2}-0.6$ (we need to convert the angle from w.r.t the normal of the wall to w.r.t the forward direction of the robot). The horizontal line cut through different regions of the diagram, representing the possible edges that the robot can bounce to at different points on the boundary with bounce angle $\theta_0$. For example, at edge $v_5v_0$, the robot can bounce to edge $v_0v_1$ and $v_1v_2$. We can create a graph for $\theta = \theta_0$, whose nodes are edges in $P'$ and two nodes are connected by a directed edge if the robot can bounce from one edge to another at $\theta_0$, as shown in Fig~\ref{fig:limit_cycle_diagram}. It is more stable for the robot to bounce to some edges than others. In the diagram (Fig~\ref{fig:concave_quad_bvd}), the line $\theta = \theta_0$ cuts through four different regions for $x \in (2, 3)$, but the sections for $v_4v_5$ and $v_0v_5$ are much larger than those for $v_0v_1$ and $v_1v_2$. So we know it is more likely and more stable for the robot to bounce to edge $v_4v_5$ and $v_0v_5$. We marked the edge with unstable bounce in the graph with $\omega$ in Fig~\ref{fig:limit_cycle_diagram}. The limit cycle shown in Fig~\ref{fig:limit_cycle} corresponds to the cycle $v_0v_1\rightarrow v_3v_4\rightarrow v_4v_5 \rightarrow v_1v_2 \rightarrow v_2v_3 \rightarrow v_0v_5 \rightarrow v_0v_1$ in Fig~\ref{fig:limit_cycle_diagram}.
(\textcolor{red}{do we talk about this here or later in the task})
\begin{figure}
\centering
\makebox[\linewidth][c]{%
\subfigure[]{
\centering
\includegraphics[width=0.45\linewidth]{images/limit_cycle_concave_quad.png}
\label{fig:limit_cycle}
}
\subfigure[]{
\centering
\includegraphics[width=0.52\linewidth]{images/concave_quad.png}
}
\label{fig:color_concave_quad}
}\\
\subfigure[]{
\centering
\includegraphics[width=0.8\linewidth]{images/limit_cycle_bvd.png}
\label{fig:concave_quad_bvd}
}
\subfigure[]{
\centering
\includegraphics[width=0.8\linewidth]{images/limit_cycle_graph.png}
\label{fig:limit_cycle_diagram}
}
\caption{The bounce visibility diagram with limit cycle information}
\label{fig:limit_cycle_bvd}
\end{figure}

@@

We might hope that if $r$ is large, then not all of the reflex vertices will produce a large number of new vertices, or that if there are many new vertices, then the new vertices will not have good visibility in $P'$ since most of them will be inserted into a corner of the polygon, and then we can bound the complexity of the BVD. Unfortunately, the number of reflex vertices, the new vertices produced in their partial local sequence, and the new vertices' visibility can be large at the same time. We will present a family of input polygons with BVD of $\Theta(n^4)$ complexity as follows. 
\begin{figure}
\hspace{-40px}
\centering
\makebox[\linewidth][c]{%
\hspace{30px}
\subfigure[]{
\centering
\includegraphics[width=0.7\linewidth]{images/chestnut_5.png}
\label{fig:t4}
}
\hspace{-40px}

\subfigure[]{
\centering
\includegraphics[width=0.7\linewidth]{images/chestnut_7.png}
}
\label{fig:t6}
}\\
\caption{The worse case polygon for $t = 4$ and $t = 6$}
\label{fig:chestnut}
\end{figure}

>>>>

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage[font=footnotesize,labelfont=footnotesize]{caption}
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
%\usepackage{begin{subfigure} % use for side-by-side figures
\usepackage{subcaption}
\usepackage{hyperref} % use for hypertext links, including those to external documents and URLs
\usepackage{multirow}
\usepackage{rotating}
\usepackage{array}
\usepackage{xspace}
\usepackage{indentfirst}
\usepackage{amssymb}
\usepackage{float}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{comment}
\usepackage{sidecap}
\usepackage{booktabs}

\title{A Visibility-Based Combinatorial Data Structure for
Mobile Robots in Polygons}

@@

The \textit{vertex visibility graph} $G_V(P) = (V, E)$ has $V = \{v_0, v_1, \ldots, v_{n-1}\}$ and $v_iv_j\in E$ if and only if the segment connecting $v_i$ and $v_j$ is completely contained in the interior of $P$. We call $v_i, v_j$ visible to each other if $v_iv_j\in E$. By definition, the vertex visibility graph for a convex polygon with $n$ vertices is a $n-$clique. Fig
\ref{fig:vvg} shows the visibility graph for a non-convex polygon.

\begin{figure}
    \includegraphics[width=0.8\linewidth]{images/viz_graph.png}
    \centering
    \caption{The vertex visibility graph}\label{fig:vvg}
    \centering
\end{figure}

We can also define the \textit{vertex-edge visibility graph} for a polygon. Using the definition in \cite{rourke_viz}, the \textit{vertex-edge visibility graph} for $P$, $G_E(P) = (V, E)$, is a bipartite graph where $V = \{\text{all vertices and edges in $P$}\}$ and $v_ie_j\in E$ if and only if $v_i$ can see the interior of $e_j$, i.e., $\exists$ point $q$ on the open edge $e_j$ such that $v_i$ and $q$ are visible. \cite{rourke_viz} established that the vertex-edge visibility graph is equivalent to the edge-edge visibility graph.
Fig \ref{fig:veg} shows an example for vertex-edge visibility graph.

\begin{figure}
\centering
\begin{subfigure}{0.25\textwidth}
  \includegraphics[width=0.6\linewidth]{images/concave_pent.png}
  \label{fig:c_p}
\end{subfigure}
\begin{subfigure}{0.25\textwidth}
  \includegraphics[width=0.8\linewidth]{images/viz_edge_graph.png}
\end{subfigure}
\caption{The vertex-edge visibility graph for the polygon in Fig
\ref{fig:c_p}\label{fig:veg}}

@@

The partial local sequence, defined in \cite{rourke_viz}, 
for a vertex $v_0$ in polygon $P$ is constructed 
by shooting a ray through $v_0$ from every vertex in $P$ that is visible from
$v_0$.

See Fig \ref{fig:pls} for an example of the vertices in the partial local
sequence of $v_0$ (blue dots with accompanying rays).

\begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{images/partial_local_sequence.png}
    \caption{Partial local sequence for $v_0$}\label{fig:pls}
    \centering
\end{figure}

The partial local sequences for vertices in a polygon is helpful for defining
the equivalence classes of edge visibility - each point in each partial local
sequence defines a \emph{transition event} where the visibility polygon of the
robot gains or loses edges if the robot were sliding along the boundary.

@@

\section{Bounce Visibility Diagram}
We will now define the central structure for this paper, the
\textit{Bounce Visibility Diagram} for polygon $P$. We will first insert all new
vertices from the partial local sequence for all vertices into the
polygon $P$, creating the new polygon $P'$ with $m$ vertices. Note that $P'$ is
no longer a polygon in general position.

The $x-$axis of the diagram parameterizes the boundary of the new polygon,
denoted as $\partial P'$, in counterclockwise order to $[0, m+1)$. The
enpoints of this interval are identified, though for visualization purposes we
do not represent this explicitly. We map every
point in $\partial P'$ to a point in $[0, m+1)$; vertex $v_i$ on $\partial P'$
will map to $x = i$; all points in the open edge $v_iv_{i+1}$ will uniformly map
to $(i, i+1)$. Let $f: [0, m+1)\mapsto \partial P'$ denote this
parameterization. We will generally use the variable $s$ to refer to points on
the boundary of $\partial P'$.

The vertical axis $\theta$ of the diagram has range $[0, \pi]$, representing all
the possible departure angles of the robot from point $s \in \partial P'$. 
At angle $0$, the robot is performing ``wall following": a bounce
to the next counterclockwise vertex.
Let the robot's ``forward direction'' be the direction of the edge that the
robot is on in counterclockwise order.

For each vertex $v_i$, we can define a function $B_i: [0, n) \mapsto [0, \pi]$
for the angle between the robot's forward direction and its direction to vertex
$v_i$ when the robot stands at $f(x)\in \partial P'$. $B_i$ are piecewise
continuous functions; the discontinuities occur when $x$ is an integer, or, in
other words, at the vertices of $\partial P'$. The original vertices in $P$ mark
discontinuities because the angle reference frame for the robot changes when it
moves to a new edge; the new vertices in $P'$ mark discontinuities because the
robot has different vertex visibility view around those vertices, which
motivates us to compute the partial local sequences in the first place.
Fig \ref{fig:bvd} shows an example of the bounce visibility diagram.

\begin{figure}

\centering
\begin{subfigure}{0.25\textwidth}
\centering
\includegraphics[width=0.7\linewidth]{images/color_pent.png}
\label{fig:color_pent}
\end{subfigure}%
\begin{subfigure}{0.25\textwidth}
\centering
\includegraphics[width=0.8\linewidth]{images/bvd.png}
\end{subfigure}
\caption{The bounce visibility diagram for the polygon shown in Fig~\ref{fig:color_pent}. }
\label{fig:bvd}

@@

\begin{figure}
% \makebox[\linewidth][c]{%
\begin{subfigure}{0.25\textwidth}
\centering
\includegraphics[width=0.5\linewidth]{images/regular_pent.png}
\end{subfigure}
\begin{subfigure}{0.25\textwidth}
\includegraphics[width=0.8\linewidth]{images/regular_pent_bvd.png}
\end{subfigure}
\centering
% }\\
\caption{The bounce visibility diagram for the regular pentagon. }
\label{fig:regular_pent_bvd}
\end{figure}

We can easily extract information for the fixed angle bouncing strategy
mentioned in section \ref{subsec:bounce_strategy} by overlaying an horizontal
line $\theta = \theta_0$ onto the BVD, where $\theta_0$ is the angle chosen by
the bouncing strategy. In Fig \ref{fig:limit_cycle_bvd}, one of the possible
limit cycle with fixed angle $\theta = 0.6$ for the concave quadrilateral is
shown in Fig \ref{fig:limit_cycle}; the bounce visibility diagram for this
polygon is shown in Fig \ref{fig:concave_quad_bvd} with the horizontal line
$\theta = \theta_0$, where $\theta_0 = \frac{\pi}{2}-0.6$ (we need to convert
the angle from w.r.t the normal of the wall to w.r.t the forward direction of
the robot). The horizontal line cut through different regions of the diagram,
representing the possible edges that the robot can bounce to at different points
on the boundary with bounce angle $\theta_0$. For example, at edge $v_5v_0$, the
robot can bounce to edge $v_0v_1$ and $v_1v_2$. We can create a graph for
$\theta = \theta_0$, whose nodes are edges in $P'$ and two nodes are connected
by a directed edge if the robot can bounce from one edge to another at
$\theta_0$, as shown in Fig \ref{fig:limit_cycle_diagram}. It is more stable
for the robot to bounce to some edges than others. In the diagram
(Fig \ref{fig:concave_quad_bvd}), the line $\theta = \theta_0$ cuts through
four different regions for $x \in (2, 3)$, but the sections for $v_4v_5$ and
$v_0v_5$ are much larger than those for $v_0v_1$ and $v_1v_2$. So we know it is
more likely and more stable for the robot to bounce to edge $v_4v_5$ and
$v_0v_5$. We marked the edge with unstable bounce in the graph with $\omega$ in
Fig \ref{fig:limit_cycle_diagram}. The limit cycle shown in
Fig \ref{fig:limit_cycle} corresponds to the cycle
$v_0v_1\rightarrow v_3v_4\rightarrow v_4v_5 \rightarrow v_1v_2 \rightarrow v_2v_3 \rightarrow v_0v_5 \rightarrow v_0v_1$
in Fig \ref{fig:limit_cycle_diagram}.
(\textcolor{red}{do we talk about this here or later in the task})

\begin{figure}
\centering
\makebox[\linewidth][c]{%

\begin{subfigure}{0.25\textwidth}
\centering
\includegraphics[width=0.45\linewidth]{images/limit_cycle_concave_quad.png}
\captionof{figure}{\label{fig:limit_cycle}}
\end{subfigure}
\begin{subfigure}{0.25\textwidth}
\centering
\includegraphics[width=0.52\linewidth]{images/concave_quad.png}
\captionof{figure}{\label{fig:color_concave_quad}}
\end{subfigure}
}

\begin{subfigure}{0.25\textwidth}
\centering
\includegraphics[width=0.8\linewidth]{images/limit_cycle_bvd.png}
\captionof{figure}{\label{fig:concave_quad_bvd}}
\end{subfigure}%
\begin{subfigure}{0.25\textwidth}
\centering
\includegraphics[width=0.8\linewidth]{images/limit_cycle_graph.png}
\captionof{figure}{\label{fig:limit_cycle_diagram}}
\end{subfigure}
\caption{The bounce visibility diagram with limit cycle information}
\label{fig:limit_cycle_bvd}
\end{figure}

@@

We might hope that if $r$ is large, then not all of the reflex vertices will produce a large number of new vertices, or that if there are many new vertices, then the new vertices will not have good visibility in $P'$ since most of them will be inserted into a corner of the polygon, and then we can bound the complexity of the BVD. Unfortunately, the number of reflex vertices, the new vertices produced in their partial local sequence, and the new vertices' visibility can be large at the same time. We will present a family of input polygons with BVD of $\Theta(n^4)$ complexity as follows.

\begin{figure}
%\hspace{-40px}
\centering
%\makebox[\linewidth][c]{%
%\hspace{30px}
\begin{subfigure}{0.25\textwidth}
\centering
\includegraphics[width=0.7\linewidth]{images/chestnut_5.png}
\label{fig:t4}
\end{subfigure}
%\hspace{-40px}

\begin{subfigure}{0.25\textwidth}
\centering
\includegraphics[width=0.7\linewidth]{images/chestnut_7.png}
\label{fig:t6}
\end{subfigure}\\
\caption{The worse case polygon for $t = 4$ and $t = 6$ 
\label{fig:chestnut}}
\end{figure}
